---
output:
  html_document: default
  pdf_document: default
---
```{r}
library(gapminder)
library(tidyverse)
library(car)
library(dplyr)
library(e1071)

```
1. Assess normality of Life Expectancy in the Americas in the year 2007. Include all
appropriate plots and outcomes, and explain your decision of whether it is normal or
not. [3 marks]

```{r}
# Loading data and subsetting
df <- gapminder %>%
  filter (year == 2007)
df1<-df$lifeExp
length(df1)
dfg1 <- data.frame(value=df1,
                   ID=seq(1,142,by=1))

```




Plot the histogram 
```{r}
ggplot(dat=dfg1, aes(x=value))+
  geom_histogram(binwidth=0.1, alpha=0.5, position="identity")

```

Check normality using qq plot
```{r}
qqnorm(df1)
```

Check skewness
```{r}
print(skewness(df1))
```


```{r}
shapiro.test(x = df1)
```

```{r}
ks.test(x = df1,
        y = "pnorm",
        mean = mean(df1),
        sd = sd(df1))
```

From the result of q-q plot, Shapiro-Wilk normality test, and One-sample Kolmogorov-Smirnov test. The normality of this dataset is confirmed.


2. Are there ‘population’ outliers when comparing Asian countries to European countries
in 1952? What are the different options to handle outliers, and what do you think is the
best option to handle outliers when making this comparison? [3 marks]

```{r}
df <- gapminder %>%
  filter (year == 1952 & (continent == "Asia" | continent == "Europe") )

library(ggplot2)
ggplot(df , aes(x=continent, y=pop)) +
  geom_boxplot()
```
This plot shows that the outlier exist in both group. The optins to deal with outliers:
1. Remove them: This should ONLY be used when you are sure the data point is not a measured value
• E.g., Heart rate data above 250 bpm; not plausible
2. Transform the data
• E.g., log/power/inverse
3. Change the score: This may seem like cheating, but if your data point is going to bias your data anyways, this is better than leaving it in
• E.g., The next highest score + 1
• E.g., Change it to 3SD above the mean; make it physiological
4. Use non-parametrics

For this case, transform the data will be an good choice.




For the following questions, if your data breaks assumptions, you must attempt to
correct it using a data transformation. If no normality transformations help, proceed as if the assumptions are intact and state this in your response.

2. Life Expectancy: Conduct the appropriate statistical test, demonstrate the underlying
assumptions, plot the results, and briefly report the results in a ‘publishable’ format, to answer a research question of your choice phrased in the following format: [5 marks]

Is life expectancy greater in [continent] compared to [continent] in [year]?
```{r}
# Subsetting
df <- gapminder %>%
  filter (year == 2007 & (continent == "Asia" | continent == "Europe") )

dfa <- gapminder %>%
  filter (year == 2007 & continent == "Asia")

dfe <- gapminder %>%
  filter (year == 2007 & continent == "Europe")


library(ggplot2)
ggplot(df , aes(x=continent, y=lifeExp)) +
  geom_boxplot()

```

```{r}
# Visualize the subset distribution by Histogram Plot 

ggplot(df, aes(x=lifeExp, fill=continent)) +
  geom_histogram(binwidth=0.1, alpha=0.5, position="identity")
```




```{r}

# visualizing normality by q-q plot of the subset life expectancy in Asia at 2007
qqnorm(dfa$lifeExp)

```

```{r}
# Check the skewness of the subset life expectancy in Asia at 2007
print(skewness(dfa$lifeExp))

```

```{r}
# Shapiro-Wilk normality test of the subset life expectancy in Asia at 2007
shapiro.test(x = dfa$lifeExp)

```

```{r}
# One-sample Kolmogorov-Smirnov test of the subset life expectancy in Asia at 2007

ks.test(x = dfa$lifeExp,
        y = "pnorm",
        mean = mean(dfa$lifeExp),
        sd = sd(dfa$lifeExp))
```


```{r}
# visualizing normality by q-q plot of the subset life expectancy in Europe at 2007
qqnorm(dfe$lifeExp)
```


```{r}
# Check the skewness of the subset life expectancy in Europe at 2007
print(skewness(dfe$lifeExp))
```

```{r}
# Shapiro-Wilk normality test of the subset life expectancy in Europe at 2007
shapiro.test(x = dfe$lifeExp)
```
```{r}
# One-sample Kolmogorov-Smirnov test of the subset life expectancy in Europe at 2007

ks.test(x = dfe$lifeExp,
        y = "pnorm",
        mean = mean(dfe$lifeExp),
        sd = sd(dfe$lifeExp))
```


Shapiro-Wilk normality test results show that both the dataset of 
life expectancy in Europe and Asia in 2007 are non-normal. So the 
data will be log-transformed.
```{r}
# Fix the outcome with a log transform 
dft <- df %>% 
  mutate(value_log = log(lifeExp))
dfat <- dft %>%
  filter (year == 2007 & continent == "Asia")
dfet <- dft %>%
  filter (year == 2007 & continent == "Europe")

```

Run normality check again on the subset life expectancy in Asia at 2007
```{r}
qqnorm(dfat$value_log)
print(skewness(dfat$value_log))
shapiro.test(x = dfat$value_log)
ks.test(x = dfat$value_log,
        y = "pnorm",
        mean = mean(dfat$value_log),
        sd = sd(dfat$value_log))
```

Run normality again on the subset life expectancy in Europe at 2007
```{r}
qqnorm(dfet$value_log)
print(skewness(dfet$value_log))
shapiro.test(x = dfet$value_log)
ks.test(x = dfet$value_log,
        y = "pnorm",
        mean = mean(dfet$value_log),
        sd = sd(dfet$value_log))
```

After log-transform the subsets, neither subset is normal.


```{r}
#Check homogeneity of variance using Levene's Test
leveneTest(lifeExp~continent, data = dft, center = mean)

```

```{r}
ggplot(dat = dft, aes(x = continent, y = lifeExp))+
  geom_violin()+
  geom_point()
```


```{r}
# Independent T-test 
t.test(x = dfat$value_log,
       y = dfet$value_log,
       alternative = "two.sided",
       var.equal = FALSE,
       paired = FALSE)
```
An independent t-test will be used to analyze differences between
conditions. The Kolmogorov-Smirnov test will be used to test for normality, and
Levene’s test for homogeneity of variance. If data is found to be non-normal, data
will be log-transformed. Significance will be set at α < 0.05.

Subject characteristics are presented as above. Life expectancy in Europe at 2007 (p = 0.02119) and life expectancy in Asia at 2007 (p = 0.03274) were found to be significantly non-normal, and were log-transformed. Using the log-transformed data, both subsets are still non-normal, and the variance was significantly different between groups (p = 0.001069). The t-test result shows that the life expectancy in Europe at 2007 is longer that that in Asia (p = 5.105e-05).



3. Wealth: Conduct the appropriate statistical test, demonstrate the underlying
assumptions, plot the results, and briefly report the results in a ‘publishable’ format, to answer a research question of your choice phased in the following format: [5 marks]
Does GDP (gdpPercap) increase between [2002] and [2007] in [Europe]?

```{r}
# Subsetting
dfw <- gapminder %>%
  filter (continent == "Europe" & (year == 2002 | year == 2007 ) )
df2 <- gapminder %>%
  filter (year == 2002 & continent == "Europe")
df7 <- gapminder %>%
  filter (year == 2007 & continent == "Europe")

# log transform
dfwt <- dfw %>% 
  mutate(value_log = log(gdpPercap))
df2t <- dfwt %>%
  filter (year == 2002 & continent == "Europe")

```


```{r}
# Check normality of GDP in Europe at 2002
qqnorm(df2$gdpPercap)
print(skewness(df2$gdpPercap))
shapiro.test(x = df2$gdpPercap)
ks.test(x = df2$gdpPercap,
        y = "pnorm",
        mean = mean(df2$gdpPercap),
        sd = sd(df2$gdpPercap))
```

```{r}
# Check normality of GDP in Europe at 2007
qqnorm(df7$gdpPercap)
print(skewness(df7$gdpPercap))
shapiro.test(x = df7$gdpPercap)
ks.test(x = df7$gdpPercap,
        y = "pnorm",
        mean = mean(df7$gdpPercap),
        sd = sd(df7$gdpPercap))
```


```{r}
# Check normalization again after dataset of GDP in Europe at 2002 is log-transformed
qqnorm(df2t$value_log)
print(skewness(df2t$value_log))
shapiro.test(x = df2t$value_log)
ks.test(x = df2t$value_log,
        y = "pnorm",
        mean = mean(df2t$value_log),
        sd = sd(df2t$value_log))
```

```{r}
# Dependent T-test 
t.test(x = df2$gdpPercap,
       y = df7$gdpPercap,
       alternative = "two.sided",
       paired = TRUE)
```
```{r}
ggplot(dat = dfw, aes(x = year, y = gdpPercap))+
  geom_point()
  
```
A dependent t-test will be used to analyze the change of GDP in Europe from 2002 to 2007. The Kolmogorov-Smirnov test will be used to test for normality, and since this is a dependent t-test, Levene’s test for homogeneity of variance will not be appliable. If data is found to be non-normal, data will be log-transformed. Significance will be set at α < 0.05.

Subject characteristics are presented as above. GDP in Europe at 2002 (p = 0.0275) was found to be significantly non-normal, and were log-transformed. Using the log-transformed data, GDP in Europe at 2002 is still non-normal (p=0.002271), the dependent t-test result shows that GDP in Europe from at 2007 is hight that that in 2002 (p=4.627e-13).








