
1. Gus T. Wind, the breathing coach, is examining the reliability of a new gas
analyzer to measure the relative concentrations of oxygen and carbon dioxide
during maximal exercise tests. Volunteers breathed into the same machine six
different times and the percent oxygen was recorded. Ideally, Dr. Wind would like
athletes to breathe only once after competitions to save time. (outcomes: gas1-
gas6)
    Intra-rater, Two-way Mix, Absolute Agreement Type
    ICC(3,1), Single Measure

2. Cy Kyle Fast, the cycling coach, is interested in whether any person would be
reliable in their quickness changing a new easy-latch bike tire in case of flats. Six
of his team members each changed 27 tires, although 2 of his assistants only
had 1 hand due to unfortunate biking accidents, and were a bit slower than the
others. (outcomes: tired1-tired6)
    Inter-rater Reliability, Two- way Random, Consistency Type
    ICC(2,27), Average Measure 

3. Juan R. Em has a new device for measuring muscle power but the machine still
needs a person to manually identify breakpoints on print outs. To measure his
own reliability, Dr. Em examines data from 27 athletes and measures the
breakpoints 6 different times. It is very important that he identifies the real power
value to set training schedules for athletes, so he would plan to take the average
of any repeated measurements. (outcomes: power1-power6)
    Test-retest, Two-Way Mix, Absolute Agreement Type
    ICC(3,6), Average Measures

4. In addition to the reliability testing, Dr. Wind was also interested in the validity of
his new gas analyzer and compared it to a top-of-the-line commercial model
(outcomes: NewGas and TopGas). He produced the following Bland-Altman plot:
a) What conclusion can you draw about the validity of the new machine? [2
marks]

As shown in the Bland-Altman plot, the new machine has the approximately equal validity to the top-of-the-line commercial model.


b) If Dr. Wind was attempted to measure the percent oxygen within a 5 unit
limit, would your conclusions change? Why? [1 mark]

the new machine has different validity comparing with the top-of-the-line commercial model. Because most measure ponits fall out of the 5 unit limit lines.


```{r}
library(tidyverse)
library(psych)

df <- read.csv("A6_Reliability.csv")

```

```{r}
##Bland-Altman 
library(BlandAltmanLeh)

#Create BA plot
bland.altman.plot(group1 = df$NewGas, group2 = df$TopGas, graph.sys = "ggplot2")
```


c) Plot NewGas against TopGas, as well as a diagonal line representing a
perfect r=1 relationship (i.e., the identity line). How does this plot compare
to the Bland-Altman plot? Why might one be better than the other? [3
marks]

Bland and Altman introduced the Bland-Altman plot to describe agreement between two quantitative measurements by constructing limits of agreement. These statistical limits are calculated by using the mean and the standard deviation (s) of the differences between two measurements. The graph displays a scatter diagram of the differences plotted against the averages of the two measurements. Horizontal lines are drawn at the mean difference, and at the limits of agreement. The limits of agreement are defined as the mean difference ± 1.96 SD of differences. If these limits do not exceed the maximum allowed difference between methods Δ (the differences within mean ± 1.96 SD are not clinically important), the two methods are considered to be in agreement and may be used interchangeably. On the other hand, identity line can be useful, but not as much information as BA plot. 


```{r}
#Identify line plot
ggplot(data = df, aes(x = NewGas, y = TopGas))+
  geom_point()+
  geom_abline(slope = 1, intercept = 0)+
  xlim(0,60)+
  ylim(0,60)
```



