---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---


1. Read in .csv data from the github address into a data frame (call it ‘bg’)
```{r}
library (RCurl)
download <- getURL("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-12/board_games.csv")
bg <- read.csv (text = download)
```


print the structure of the data
```{r}
str(bg)
```
Appreciate the scope of the dataset:
  use ggplot2 to create a geom_point plot 
```{r}
library(ggplot2)
ggplot(bg, aes(x=year_published, y=average_rating)) +
  geom_point(aes(colour = users_rated), size =0.1, alpha = 0.1)+
scale_colour_gradient(low = "red", high = "white")

```
  use ggplot2 to create a geom_jitter plot 
```{r}
ggplot(bg, aes(x=year_published, y=average_rating)) +
  geom_jitter(aes(colour = users_rated), size =0.1, alpha = 0.1)+
  scale_colour_gradient(low = "red", high = "white")

```
Display both plots together side-by-side using the grid.arrange function from the gridExtra package. 

What is the difference between plots and when would one be more appropriate?

    Jittering is adding a small amount of random noise to data. It is often used to spread out points that would otherwise be overplotted. It is only effective in the non-continuous data case where overplotted points typically are surrounded by whitespace - jittering the data into the whitespace allows the individual points to be seen. It effectively un-discretizes the discrete data.
    
    With high density data, jittering doesn't help because there is not a reliable area of whitespace around overlapping points. Other common techniques for mitigating overplotting include
    
    using smaller points
    using transparency
    binning data (as in a heat map)


```{r}
# install.packages("gridExtra")

require(gridExtra)

p1 <- ggplot(bg, aes(x=year_published, y=average_rating)) +
  geom_point(aes(colour = users_rated), size =0.1, alpha = 0.1)+
  scale_colour_gradient(low = "red", high = "white")


p2 <- ggplot(bg, aes(x=year_published, y=average_rating)) +
  geom_jitter(aes(colour = users_rated), size =0.1, alpha = 0.1)+
  scale_colour_gradient(low = "red", high = "white")
grid.arrange(p1, p2, nrow = 1)
```
```{r}
library(dplyr)
library(stringr)
bg2 <- bg %>%
  filter(str_detect(category, paste(c("Card Game", "Dice"),collapse = '|')))

bg3<-mutate(bg2, type = 
         ifelse((!grepl("Card Game", category)), "Dice",
          ifelse((!grepl("Dice", category)), "Card Game", "Both")))
```

```{r}
library(ggplot2)
ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_jitter()
  
```


```{r}
library(ggplot2)
ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_boxplot()
```



```{r}
ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_violin()
```


```{r}
require(gridExtra)

p3<-ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_jitter()

p4<-ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_boxplot()
  
p5<-ggplot(bg3, aes(x=type, y=average_rating)) +
  geom_violin()

grid.arrange(p3, p4, p5, nrow = 1)
```
In a few sentences, describe how each type of plot provides unique information about the dataset. Is one superior to the others?
            
        The jitter geom is a convenient shortcut for geom_point(position = "jitter"). It adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets.
        
        The boxplot compactly displays the distribution of a continuous variable. It visualises five summary statistics (the median, two hinges and two whiskers), and all "outlying" points individually.    
        
        A violin plot is a compact display of a continuous distribution. It is a blend of geom_boxplot() and geom_density(): a violin plot is a mirrored density plot displayed in the same way as a boxplot.
        
        Different type of plots used in the different situations depending on the request. There is no one superior to the others.



Are games published later than year 1995 ranked higher than games earlier than year 1995?

```{r}
bg4<-mutate(bg, year_1995 = ifelse(year_published <= 1995, "earlier_than_year_1995",
                                     "later_than_year_1995"))
```


```{r}
library(ggplot2)
ggplot(bg4, aes(x=year_1995, y=average_rating)) +
  geom_boxplot(col=topo.colors(1))+
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 1))+
  ggtitle("Games published year affects its rating") +
  theme(plot.title = element_text(hjust=0, size=16)) +
  xlab("year")+
  ylab("average rating")+
  labs(
       subtitle = "Are games published later than year 1995 ranked higher than games earlier than year 1995 ?",
       caption = "Dataset: Board game rankings from boardgamegeek.com") 
  
 
  
```







