---
title: "R Notebook Example"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

This is an R Markdown Notebook. When you execute code within the notebook, the results appear beneath the code. For your assignments, I will expect you to submit a .Rmd file so you can have written answers to the assignment questions, with modifiable code so myself and the TA will be able to examine any errors

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

For every 'custom' package we use, we have to re-load the library into R during each new session using the library() function. First, let's simulate some data and look at simple descriptive statistics:

```{r}
library(tidyverse)
g1 <- rnorm(n=100, mean=350, sd=20)     #Creates a random sample with mean 350 and SD 20
g2 <- rnorm(n=100, mean=280, sd=8)      #Creates a random sample with mean 280 and SD 8
df <- data.frame(g1,g2) %>%             #Creates a 'wide' format data.frame, then uses the Tidyverse Pipeline to call next function
  gather(key="group",value="value")     #Converts into a 'long' format data.frame

```

Nothing happened? That's because none of the commands told R to print any data. We can tell it to show us some numbers by using the handy tapply function from Tidyverse. This applies a function for a specific variable, in certain conditions. In this example, we want to see the summary statistics of our value for each group:

```{r}
tapply(df$value, df$group, summary)

```

These statistics show us basic information about our data, like the interquartile range, mean, min and max. I prefer to see the visual representation of the data, so let's take a look at the histogram of the data we've generated:

```{r}
ggplot(dat=df, aes(x=value,fill=group)) +
  geom_histogram(binwidth=10,alpha=0.5,position="identity")
```

And that's it! For each component of your assignments, try to describe what you will do, then add the R functions underneath.
